---
title: "Class 1 Notebook"
output: html_notebook
---

We will demonstrate:
1. importing data into R from .csv files
2. generating frequencies  
3. calculating proportions
4. creating histograms
5. missing values in R

### Sets all the file locations
- Someone can come in and change the file location to their own computer
- Better at the top of your code
- one place to change so it runs for everyone (reproducible)

```{r Sets_file_locations}
# lets you a windows interface to pick the file location and store in a variable
MyFileLocation = file.choose()
MyFileLocation
```

### Importing Data
- read.csv function convert a comma separated text file to a dataframe
- do not need to specify file = ""
- many options
- use ?read.csv to learn more

```{r IMPORT DATA}
infantcarseat<-read.csv( file = MyFileLocation)
names(infantcarseat) #shows you variables in infantcarseat DataFrame
```

We can even look at the first few rows (in console)

```{r viewing first few rows of a DataFrame}
head(infantcarseat)
```

#### Notes:
In RStudio, we can just click on the DataFrame in the Environment tab and open it

### Attaching variables to global environment
- allows you to reference just the variable
- does not require you to specify the dataframe
- can cause problems if you attach >2 datasets with same variable names
- most data scientists DO NOT DO THIS.

```{r Attaching variables from infantcarseat}
attach(infantcarseat)
carseat
```

#### Notes:
- Look at window-panes on the left
- all variables are now (separately) loaded in the "Environment" tab

You can always reference a variable using dollar ($) character
- infantcarseat$carseat
- infantcarseat$sex
- infantcarseat$surfactant

For simplicity I will just use the variable names loaded into global.env

```{r Using Table to explore frequency counts}
table(carseat)
table(o2sat)
table(gage)
table(surfactant)
table(failreason)
table(sex)
```

We can also use the summary() function to explore the data

```{r Using summary function for frequency counts}
summary(sex)
summary(failreason)
summary(o2sat)
```

# Functions for Descriptive Statistics

```{r Descriptive Stats}
mean(o2sat)
sd(o2sat)
mean(gage)
sd(gage)
table(sex)
prop.table(table(sex))
table(carseat)
prop.table(table(carseat))
table(caffeine)
prop.table(table(caffeine))
table(surfactant)
prop.table(table(surfactant))
table(failreason)
prop.table(table(failreason))
```

# Visualizing Descriptive Stats and data
- ?hist
```{r Histograms}
hist(o2sat,main="Histogram of Oxygen Saturation",xlab="Oxygen Saturation")
hist(gage,main="Histogram of Gestational Age", xlab="Gestational Age")
```
```{r Boxplots}
boxplot( x = o2sat,
         main = "Oxygen saturation",
         ylab = "%conc")
```


### Missing values in R
- NA indicates "missing"
- many functions will skip NA values
- may need to use function options to use or show NA values

```{r missing values}
Vector.1 = c(1,3,5,9)
Vector.2 = c(2,4,NA,8)
VectorSum = Vector.1 + Vector.2
VectorSum
summary(VectorSum)
table(VectorSum)
table(VectorSum, useNA = "ifany")
table(failreason)
```

