---
title: "Session 3 Notebook"
output: html_notebook
---

```{r FILE LOCATIONS}
InFile_location = file.choose()
```


# Measures of Disease Frequency and Association

## Question 1a--Prevalence

In 2016, there were 21 million adults with Type 2 Diabetes out of a total of 244,186,465 adults in the US. 

(a) Calculate the prevalence of Type 2 Diabetes in adults in the US in 2016.

```{r}
21*(10^6) / 244186465
```

(b)	Interpret the prevalence you calculated in part a.

```{r}

```

## Question 1b

In 2018, a survey was conducted of 1,891 children 12-23 months of age in Zambia.  Among these children, 1,719 reported receiving the measles vaccine.

(a)	Calculate the prevalence of measles vaccination among these 12-23 month-old children in Zambia in 2018.



(b)	Interpret the prevalence you calculated in part a.



## Question 2

Who is part of the at risk population? Consider the following two examples.

(a)	Who is at risk for influenza infection?



(b)	Who is at risk for incident endometrial cancer?


## Question 3
In a cohort of pregnant women receiving prenatal care at BMC during a 1-year period from 2014-2015, there were 252 deliveries. Of these newborns, 13 were classified as low birthweight (<2,500 grams). 

(a)	Calculate the 1-year cumulative incidence (risk) of low birthweight in this group.



(b)	Interpret the 1-year cumulative incidence (risk) you calculated in part a.



(c)	Compute the 1-year cumulative incidence (risk) in R using the following code.

```{r Cumulative Incidence}
DrugUse <- read.csv( file = InFile_location)

DrugUse$LowBirthWt <- with(DrugUse,
                           ifelse (BirthWt <2500, 1, 0)
)

with(DrugUse, 
     table(DrugUse$LowBirthWt)
)

with(DrugUse,
     prop.table( table(LowBirthWt) )
)

```

(d)	Suppose we want to estimate the cumulative incidence (risk) of low birthweight in first pregnancies. Of the 252 deliveries, 67.1% of infants were a first pregnancy. Of these newborns, 11 were classified as low birthweight. Calculate the one-year cumulative incidence (risk) of low birthweight in this group.


## Question 4
Let’s discuss issues with tracing people over a longer period of time.

(a)	What are some potential problems? What might happen to people?


(b)	Can you think of some populations where a long follow-up period might be especially problematic?


## Question 5
In an open cohort of children under 5 in the UK, the effect of socio-economic deprivation on injury was examined from 1990-2009. A total of 20,668 fractures, 15,796 burns and 10,155 poisonings were reported over a total of 2,725,000 person-years. 

(a)	Calculate the incidence rate of burns among children under 5 in the UK from 1990-2009.



(b)	Interpret the incidence rate you calculated in part a.


## Question 6
In a cohort of pregnant women receiving prenatal care at BMC from 2014-2015, there were 252 deliveries. Of these newborns, 13 were classified as low birthweight (<2,500 grams). Among 130 pregnant women who ever used cocaine during pregnancy, there were 9 newborns with low birthweight. 

(a)	Calculate the risk difference (absolute effect) for ever using cocaine during pregnancy, compared to never, on the incidence of low birthweight at BMC from 2014-2015.


(b)	Interpret the risk difference you calculated in part a.


